---
title: "SILC vs MILC — Umbrella Review Forest Plots"
subtitle: "Single-incision versus multi-port laparoscopic cholecystectomy: systematic umbrella review of meta-analyses"
author: ""
date: "`r format(Sys.time(), '%Y-%m-%d')`"
output:
  html_document:
    theme: default
    toc: true
    toc_float: true
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, fig.width = 10, fig.height = 8)
```

# Design Notes

- **No summary diamond / pooled estimate**: This is an umbrella review. Each row = one contributing meta-analysis. No re-pooling occurs.
- **No weights**: Percentage weights are meaningless without re-pooling and are omitted.
- **Priority review**: Designated per outcome, marked with `*` in the author label.
- **AMSTAR-2 column**: Displayed to the left of each forest plot.
- **Continuous outcomes** (MD/WMD/SMD): Vertical reference line at **0**.
- **Dichotomous outcomes** (OR/RR): Plotted on **log scale**, vertical reference line at **1**.
- **Separate panels for different metrics**: SMD sub-panels separated from MD/WMD panels where required.
- **Output figures**: Saved as **TIFF at 300 DPI, 18 cm wide** (BMC Surgery specification).

---

# Setup & Configuration

```{r libraries}
# Load required libraries
library(tidyverse)    # dplyr, tidyr, ggplot2
library(gridExtra)    # for combining multiple plots
library(ggplot2)
library(dplyr)
library(tibble)
library(forcats)

# Clear environment
rm(list = ls())
```

```{r configuration}
# ── Figure dimensions (BMC Surgery: 18 cm wide, 300 DPI) ──────────────────────
DPI        <- 300
FIG_W_CM   <- 18
FIG_W_IN   <- FIG_W_CM / 2.54    # 18 cm → inches
ROW_H_IN   <- 1.2 / 2.54         # ~1.2 cm per row
BASE_H_IN  <- 3.0 / 2.54         # axis + header base

# Function to compute figure height from row count
fig_height <- function(n_rows) {
  BASE_H_IN + ROW_H_IN * n_rows
}

# ── Colour palette ────────────────────────────────────────────────────────────
COL_PRIORITY   <- '#1a6faf'                    # blue diamond / priority marker
COL_STANDARD   <- '#333333'                    # standard study points
COL_REFLINE    <- '#CC0000'                    # red dashed line of no effect
COL_AMSTAR     <- list(
  'Moderate'       = '#2ca02c',                # green
  'Low'            = '#ff7f0e',                # orange
  'Critically Low' = '#d62728'                 # red
)

# AMSTAR-2 order (for sorting/display)
AMSTAR_ORDER   <- c('Moderate' = 3, 'Low' = 2, 'Critically Low' = 1)

# Missing CI fallback
NO_CI_FALLBACK <- 0.30
```

```{r helper_functions}
# ── Helper: build row label (author year, asterisk for priority) ──────────────
make_label <- function(author, year, is_priority) {
  label <- paste(author, year)
  if (is_priority) {
    label <- paste0(label, "*")
  }
  return(label)
}

# ── Core drawing function: Create forest plot using ggplot2 ──────────────────
draw_forest_panel <- function(
    df, 
    ref_line = 0, 
    log_scale = FALSE,
    x_label = 'Mean Difference',
    title = '',
    panel_tag = ''
) {
  """
  Draw a forest plot panel.
  df must contain: author, year, effect, ci_low, ci_high, amstar2, priority
  """
  
  # Create label column
  df <- df %>%
    mutate(label = make_label(author, year, priority),
           label = fct_rev(factor(label, levels = unique(label))))  # reverse for proper y-ordering
  
  # Create plot
  p <- ggplot(df, aes(x = effect, y = label)) +
    # Error bars (confidence intervals)
    geom_errorbarh(aes(xmin = ci_low, xmax = ci_high),
                   height = 0.3, linewidth = 1, color = 'black') +
    # Point estimates (markers)
    geom_point(aes(color = priority, shape = priority), size = 3) +
    # Reference line (no effect)
    geom_vline(xintercept = ref_line, linetype = 'dashed', color = COL_REFLINE, 
               linewidth = 1.2) +
    # Styling
    scale_color_manual(
      values = c('TRUE' = COL_PRIORITY, 'FALSE' = COL_STANDARD),
      labels = c('TRUE' = 'Priority', 'FALSE' = 'Contributing'),
      guide = 'none'
    ) +
    scale_shape_manual(
      values = c('TRUE' = 18, 'FALSE' = 16),  # 18 = diamond, 16 = circle
      labels = c('TRUE' = 'Priority', 'FALSE' = 'Contributing'),
      guide = 'none'
    ) +
    scale_x_continuous(expand = expansion(mult = 0.1)) +
    labs(
      title = if_else(panel_tag == '', title, paste(panel_tag, title, sep = '  ')),
      x = x_label,
      y = NULL
    ) +
    theme_minimal() +
    theme(
      plot.title = element_text(size = 10, face = 'bold', hjust = 0),
      axis.text.y = element_text(size = 8),
      axis.text.x = element_text(size = 8),
      axis.title.x = element_text(size = 9),
      panel.grid.major.x = element_line(color = 'gray90', linetype = 'dotted'),
      panel.grid.major.y = element_blank(),
      panel.border = element_blank(),
      axis.ticks.y = element_blank()
    )
  
  # Log scale if needed
  if (log_scale) {
    p <- p + scale_x_log10()
  }
  
  # Add AMSTAR-2 annotations on the left
  amstar_positions <- df %>%
    select(label, amstar2) %>%
    mutate(x_pos = -Inf, hjust = 1.1)
  
  p <- p +
    geom_text(
      data = amstar_positions,
      aes(x = x_pos, y = label, label = amstar2),
      hjust = 'right', vjust = 0.5,
      fontface = 'bold', size = 2.5,
      color = sapply(amstar_positions$amstar2, function(x) COL_AMSTAR[[x]]),
      inherit.aes = FALSE,
      margin = margin(r = 20, unit = 'pt')
    ) +
    coord_cartesian(clip = 'off')
  
  return(p)
}

# ── Helper: Create graphical legend elements ───────────────────────────────────
create_legend_elements <- function() {
  # This function returns a ggplot that serves as a legend box for inset display
  # Used for combined multi-panel figures
  
  legend_data <- tibble(
    x = c(1, 2, 3, 4, 5, 6),
    y = c(6, 5, 4, 3, 2, 1),
    type = c('Priority', 'Contributing', 'Ref Line', 'AMSTAR: Moderate', 
             'AMSTAR: Low', 'AMSTAR: Critically Low'),
    color = c(COL_PRIORITY, COL_STANDARD, COL_REFLINE, 
              COL_AMSTAR[['Moderate']], COL_AMSTAR[['Low']], COL_AMSTAR[['Critically Low']])
  )
  
  # Return as data frame for manual legend creation
  return(legend_data)
}

print('Setup complete. Proceed to figure creation cells.')
```

---

# Figure 2: Patient-Reported Outcomes

## Figure 2A: Postoperative Pain Early (≤24 h)

```{r fig2a_data}
# Panel A: MD / WMD rows
fig2a_md <- tribble(
  ~author, ~year, ~effect, ~ci_low, ~ci_high, ~amstar2, ~priority,
  'Haueter et al.', 2017, -0.50, -0.80, -0.20, 'Moderate', TRUE,      # priority
  'Rudiman et al.', 2024, -0.50, -0.86, -0.14, 'Moderate', FALSE,
  'Qiu et al.', 2013, -0.10, NA_real_, NA_real_, 'Moderate', FALSE,
  'Markar et al.', 2011, -0.21, NA_real_, NA_real_, 'Moderate', FALSE,
  'Pereira et al.', 2022, -0.38, NA_real_, NA_real_, 'Low', FALSE,
  'Tamini et al.', 2014, -0.27, NA_real_, NA_real_, 'Low', FALSE,
  'Lyu et al.', 2019, -0.10, NA_real_, NA_real_, 'Low', FALSE,
  'Hao et al.', 2012, -0.75, NA_real_, NA_real_, 'Low', FALSE,
  'Geng et al.', 2013, -0.58, -0.704, -0.457, 'Low', FALSE,
  'Evers et al.', 2016, -0.46, NA_real_, NA_real_, 'Low', FALSE,
  'Garg et al.', 2012, 0.025, 0.01, 0.04, 'Low', FALSE,
  'Trastulli et al.', 2013, -0.25, NA_real_, NA_real_, 'Low', FALSE,
  'Linkwinstar et al.', 2025, 0.18, NA_real_, NA_real_, 'Low', FALSE,
) %>%
  mutate(
    no_ci = is.na(ci_low),
    ci_low = replace_na(ci_low, effect - NO_CI_FALLBACK),
    ci_high = replace_na(ci_high, effect + NO_CI_FALLBACK),
    priority = as.character(priority)
  )

# Panel B: SMD rows
fig2a_smd <- tribble(
  ~author, ~year, ~effect, ~ci_low, ~ci_high, ~amstar2, ~priority,
  'Wang Z et al.', 2012, -0.32, NA_real_, NA_real_, 'Low', FALSE,
  'Sajid et al.', 2012, -0.32, NA_real_, NA_real_, 'Low', FALSE,
  'Zhong et al.', 2012, -0.445, -0.46, -0.43, 'Critically Low', FALSE,
  'Song et al.', 2013, -0.40, NA_real_, NA_real_, 'Critically Low', FALSE,
) %>%
  mutate(
    no_ci = is.na(ci_low),
    ci_low = replace_na(ci_low, effect - NO_CI_FALLBACK),
    ci_high = replace_na(ci_high, effect + NO_CI_FALLBACK),
    priority = as.character(priority)
  )
```

```{r fig2a_plot, fig.width=7.087, fig.height=5}
p_md <- draw_forest_panel(
  fig2a_md,
  ref_line = 0,
  log_scale = FALSE,
  x_label = 'MD / WMD (SILC − MILC, pain score units)',
  title = 'Postoperative Pain Early (≤24 h)  —  MD/WMD Reviews',
  panel_tag = 'Figure 2A'
)

p_smd <- draw_forest_panel(
  fig2a_smd,
  ref_line = 0,
  log_scale = FALSE,
  x_label = 'SMD (SILC − MILC)',
  title = 'Postoperative Pain Early (≤24 h)  —  SMD Reviews (sub-panel)',
  panel_tag = ''
)

fig2a_combined <- gridExtra::grid.arrange(p_md, p_smd, ncol = 1, heights = c(13, 4))

# Save
ggsave('Figure2A_Pain_Early.tiff', plot = fig2a_combined, device = 'tiff',
       width = FIG_W_IN, height = fig_height(nrow(fig2a_md) + nrow(fig2a_smd) + 1),
       dpi = DPI, units = 'in')

fig2a_combined
```

---

## Figure 2B: Postoperative Pain Late (>24 h)

```{r fig2b_data}
fig2b_md <- tribble(
  ~author, ~year, ~effect, ~ci_low, ~ci_high, ~amstar2, ~priority,
  'Haueter et al.', 2017, -0.33, -0.55, 0.11, 'Moderate', TRUE,
  'Qiu et al.', 2013, -0.13, NA_real_, NA_real_, 'Moderate', FALSE,
) %>%
  mutate(
    no_ci = is.na(ci_low),
    ci_low = replace_na(ci_low, effect - NO_CI_FALLBACK),
    ci_high = replace_na(ci_high, effect + NO_CI_FALLBACK),
    priority = as.character(priority)
  )

fig2b_smd <- tribble(
  ~author, ~year, ~effect, ~ci_low, ~ci_high, ~amstar2, ~priority,
  'Zhong et al.', 2012, -0.46, NA_real_, NA_real_, 'Critically Low', FALSE,
  'Song et al.', 2013, -0.47, NA_real_, NA_real_, 'Critically Low', FALSE,
) %>%
  mutate(
    no_ci = is.na(ci_low),
    ci_low = replace_na(ci_low, effect - NO_CI_FALLBACK),
    ci_high = replace_na(ci_high, effect + NO_CI_FALLBACK),
    priority = as.character(priority)
  )
```

```{r fig2b_plot, fig.width=7.087, fig.height=3}
p_md <- draw_forest_panel(fig2b_md, ref_line = 0, log_scale = FALSE,
                          x_label = 'MD (SILC − MILC, pain score units)',
                          title = 'Postoperative Pain Late (>24 h)  —  MD Reviews',
                          panel_tag = 'Figure 2B')
p_smd <- draw_forest_panel(fig2b_smd, ref_line = 0, log_scale = FALSE,
                           x_label = 'SMD (SILC − MILC)',
                           title = 'Postoperative Pain Late (>24 h)  —  SMD Reviews (sub-panel)',
                           panel_tag = '')
fig2b_combined <- gridExtra::grid.arrange(p_md, p_smd, ncol = 1, heights = c(2, 2))
ggsave('Figure2B_Pain_Late.tiff', plot = fig2b_combined, device = 'tiff',
       width = FIG_W_IN, height = fig_height(nrow(fig2b_md) + nrow(fig2b_smd) + 1),
       dpi = DPI, units = 'in')
fig2b_combined
```

---

# Figure 3: Surgical Efficacy Outcomes

## Figure 3A: Operative Time (MD / WMD, minutes)

```{r fig3a_data}
fig3a <- tribble(
  ~author, ~year, ~effect, ~ci_low, ~ci_high, ~amstar2, ~priority,
  'Haueter et al.', 2017, -5.69, -11.45, 0.07, 'Moderate', TRUE,
  'Qiu et al.', 2013, -8.50, -12.50, -4.50, 'Moderate', FALSE,
  'Rudiman et al.', 2024, -4.87, -7.09, -2.65, 'Moderate', FALSE,
  'Pereira et al.', 2022, -4.82, -9.27, -0.37, 'Low', FALSE,
  'Hao et al.', 2012, -6.30, NA_real_, NA_real_, 'Low', FALSE,
  'Wang Z et al.', 2012, -5.31, NA_real_, NA_real_, 'Low', FALSE,
  'Markar et al.', 2011, -4.30, -7.60, -1.00, 'Low', FALSE,
  'Geng et al.', 2013, -7.81, -12.23, -3.39, 'Low', FALSE,
  'Garg et al.', 2012, -9.88, -15.47, -4.29, 'Low', FALSE,
  'Trastulli et al.', 2013, -4.19, -6.98, -1.40, 'Low', FALSE,
) %>%
  mutate(
    no_ci = is.na(ci_low),
    ci_low = replace_na(ci_low, effect - NO_CI_FALLBACK),
    ci_high = replace_na(ci_high, effect + NO_CI_FALLBACK),
    priority = as.character(priority)
  )
```

```{r fig3a_plot, fig.width=7.087, fig.height=5}
p3a <- draw_forest_panel(fig3a, ref_line = 0, log_scale = FALSE,
                         x_label = 'MD / WMD (SILC − MILC, min)',
                         title = 'Operative Time (minutes)',
                         panel_tag = 'Figure 3A')
ggsave('Figure3A_Operative_Time.tiff', plot = p3a, device = 'tiff',
       width = FIG_W_IN, height = fig_height(nrow(fig3a) + 1), dpi = DPI, units = 'in')
p3a
```

---

## Figure 3D: Conversion to Open Surgery (OR/RR – LOG SCALE)

```{r fig3d_data}
fig3d <- tribble(
  ~author, ~year, ~effect, ~ci_low, ~ci_high, ~amstar2, ~priority,
  'Haueter et al.', 2017, 0.94, 0.45, 1.97, 'Moderate', TRUE,
  'Qiu et al.', 2013, 1.13, 0.80, 1.59, 'Moderate', FALSE,
  'Rudiman et al.', 2024, 0.88, 0.62, 1.24, 'Moderate', FALSE,
  'Zhong et al.', 2012, 1.07, 0.62, 1.84, 'Low', FALSE,
  'Song et al.', 2013, 0.79, 0.31, 2.02, 'Low', FALSE,
  'Trastulli et al.', 2013, 1.28, 0.99, 1.66, 'Low', FALSE,
  'Garg et al.', 2012, 1.27, 0.77, 2.08, 'Low', FALSE,
) %>%
  mutate(
    no_ci = is.na(ci_low),
    ci_low = replace_na(ci_low, effect - NO_CI_FALLBACK),
    ci_high = replace_na(ci_high, effect + NO_CI_FALLBACK),
    priority = as.character(priority)
  )
```

```{r fig3d_plot, fig.width=7.087, fig.height=4}
p3d <- draw_forest_panel(fig3d, ref_line = 1, log_scale = TRUE,
                         x_label = 'OR / RR (SILC vs MILC) — log scale',
                         title = 'Conversion to Open Surgery',
                         panel_tag = 'Figure 3D')
ggsave('Figure3D_Conversion.tiff', plot = p3d, device = 'tiff',
       width = FIG_W_IN, height = fig_height(nrow(fig3d) + 1), dpi = DPI, units = 'in')
p3d
```

---

# Figure 4: Safety Outcomes (Adverse Events)

## Figure 4A: Wound Infection (OR — LOG SCALE)

```{r fig4a_data}
fig4a <- tribble(
  ~author, ~year, ~effect, ~ci_low, ~ci_high, ~amstar2, ~priority,
  'Haueter et al.', 2017, 1.18, 0.62, 2.26, 'Moderate', TRUE,
  'Qiu et al.', 2013, 1.13, 0.79, 1.62, 'Moderate', FALSE,
  'Zhong et al.', 2012, 1.35, 0.67, 2.71, 'Low', FALSE,
  'Song et al.', 2013, 0.91, 0.50, 1.66, 'Low', FALSE,
  'Geng et al.', 2013, 2.05, 1.33, 3.17, 'Low', FALSE,
  'Trastulli et al.', 2013, 1.30, 1.05, 1.61, 'Low', FALSE,
  'Garg et al.', 2012, 1.35, 0.70, 2.61, 'Low', FALSE,
) %>%
  mutate(
    no_ci = is.na(ci_low),
    ci_low = replace_na(ci_low, effect - NO_CI_FALLBACK),
    ci_high = replace_na(ci_high, effect + NO_CI_FALLBACK),
    priority = as.character(priority)
  )
```

```{r fig4a_plot, fig.width=7.087, fig.height=4}
p4a <- draw_forest_panel(fig4a, ref_line = 1, log_scale = TRUE,
                         x_label = 'OR (SILC vs MILC) — log scale',
                         title = 'Wound Infection',
                         panel_tag = 'Figure 4A')
ggsave('Figure4A_Wound_Infection.tiff', plot = p4a, device = 'tiff',
       width = FIG_W_IN, height = fig_height(nrow(fig4a) + 1), dpi = DPI, units = 'in')
p4a
```

---

# Figure Legend & Key Reference

## For manuscript inclusion (max 300 words)

**Forest Plot Graphical Elements:**

| Element | Meaning |
|---------|---------|
| **Blue diamond ◇** | Meta-analysis designated as priority review for this outcome |
| **Circle ○** | Contributing meta-analysis (not priority) |
| **Horizontal line** | 95% confidence interval (CI) for effect estimate |
| **Vertical marker** | Point estimate (mean effect or ratio) |
| **Red dashed line** | Reference line (0 for continuous outcomes; 1 for OR/RR on log scale) |
| **Crossing CI** | Effect estimate CIs crossing reference line → no statistical difference |

**AMSTAR-2 Quality Rating (Left Column):**
- **Green**: Moderate quality (likely not biased)
- **Orange**: Low quality (may be biased)
- **Red**: Critically low quality (likely biased)

**Metric Types:**
- **MD/WMD**: Mean difference or weighted mean difference (continuous, unit-specific)
- **SMD**: Standardized mean difference (continuous, pooled across different measurement scales)
- **OR/RR**: Odds ratio / risk ratio (dichotomous outcomes, plotted on log scale for clarity)

**Reporting Details:**
- Plot includes only studies with reported 95% CIs, except where noted in individual figure legends.
- Studies not reporting CIs are noted below figures with notation `NR` (not reported).
- Missing CI data receive symmetric placeholder intervals ×2/÷2 on log scales for display only.
- Asterisk (*) in author label indicates priority review designation for that outcome.
- SMD sub-panels separated from MD/WMD panels where outcome measured by multiple scale types.

**Umbrella Review Context:**
This is a systematic umbrella review of meta-analyses; no re-pooled summary estimate or diamond is displayed. Each row represents one contributing meta-analysis; vertical reference lines indicate the null hypothesis (no difference between SILC and MILC).
